---
title: "block_calculations"
author: "Zach Tefertiller"
date: "2025-06-25"
output: html_document
---

```{r}
# Assume df is your dataframe
colors <- c("b", "o", "y")
block_size <- 10

# Function to calculate block statistics per color
calculate_block_stats <- function(df, color_label, block_num) {
  block_name <- paste0(color_label, block_num)
  
  df %>%
    group_by(participant_id) %>%
    filter(balloon_color == color_label) %>%
    mutate(color_trial_order = row_number()) %>%
    filter(color_trial_order > (block_num - 1) * block_size &
           color_trial_order <= block_num * block_size) %>%
    summarise(
      !!paste0(block_name, "_avg") := mean(inflations, na.rm = TRUE),
      !!paste0(block_name, "_sd") := sd(inflations, na.rm = TRUE),
      !!paste0(block_name, "_adj") := mean(inflations[popped == 0], na.rm = TRUE),
      !!paste0(block_name, "_earnings") := mean(inflations[popped == 0], na.rm = TRUE) * 0.003,
      .groups = "drop"
    )
}

# Calculate blocks for each color and bind them
blocks_df <- NULL

for(color_label in colors) {
  for(block_num in 1:6) { # assuming you have up to 6 blocks per color (1-60 trials per color)
    block_stats <- calculate_block_stats(df, color_label, block_num)
    blocks_df <- if(is.null(blocks_df)) block_stats else full_join(blocks_df, block_stats, by = "participant_id")
  }
}

# Join the new data to the original dataframe (repeated values per participant)
df_extended <- left_join(df, blocks_df, by = "participant_id")
```

### Group-Level Summary

Calculate the mean and standard deviation across all participants:

```{r}
group_df <- blocks_df %>%
  summarise(across(-participant_id, list(avg_all = ~mean(.x, na.rm = TRUE), sd_all = ~sd(.x, na.rm = TRUE))))

# Append group-level data to the original dataframe (as single row repeated)
df_extended <- bind_cols(df_extended, group_df[rep(1, nrow(df_extended)), ])
```


```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
```

### Data Preparation

Replace `df` with your actual dataframe name.

```{r}
# Assume df is your dataframe
colors <- c("b", "o", "y")
block_size <- 10

# Function to calculate block statistics per color
calculate_block_stats <- function(df, color_label, block_num) {
  block_name <- paste0(color_label, block_num)
  
  df %>%
    group_by(participant_id) %>%
    filter(balloon_color == color_label) %>%
    mutate(color_trial_order = row_number()) %>%
    filter(color_trial_order > (block_num - 1) * block_size &
           color_trial_order <= block_num * block_size) %>%
    summarise(
      !!paste0(block_name, "_avg") := mean(inflations, na.rm = TRUE),
      !!paste0(block_name, "_part_sd") := sd(inflations, na.rm = TRUE),
      !!paste0(block_name, "_adj") := mean(inflations[popped == 0], na.rm = TRUE),
      !!paste0(block_name, "_earnings") := sum(inflations[popped == 0], na.rm = TRUE) * 0.003,
      .groups = "drop"
    )
}

# Calculate blocks for each color and bind them
blocks_df <- NULL

for(color_label in colors) {
  for(block_num in 1:6) { # assuming you have up to 6 blocks per color (1-60 trials per color)
    block_stats <- calculate_block_stats(df, color_label, block_num)
    blocks_df <- if(is.null(blocks_df)) block_stats else full_join(blocks_df, block_stats, by = "participant_id")
  }
}

# Join the new data to the original dataframe (repeated values per participant)
df_extended <- left_join(df, blocks_df, by = "participant_id")
```

### Group-Level Summary

Calculate the mean and standard deviation from the distribution of block averages across all participants:

```{r}
group_df <- blocks_df %>%
  summarise(across(-participant_id, 
                   list(avg_all = ~mean(.x, na.rm = TRUE), 
                        sd_all = ~sd(.x, na.rm = TRUE))))

# Append group-level data to the original dataframe (as single row repeated)
df_extended <- bind_cols(df_extended, group_df[rep(1, nrow(df_extended)), ])
```






```{r}
# Assume df is your dataframe
colors <- c("b", "o", "y")
block_size <- 10

# Function to calculate block statistics per color
calculate_block_stats <- function(df, color_label, block_num) {
  block_name <- paste0(color_label, block_num)
  
  df %>%
    group_by(participant_id) %>%
    filter(balloon_color == color_label) %>%
    mutate(color_trial_order = row_number()) %>%
    filter(color_trial_order > (block_num - 1) * block_size &
           color_trial_order <= block_num * block_size) %>%
    summarise(
      !!paste0(block_name, "_avg") := mean(inflations, na.rm = TRUE),
      !!paste0(block_name, "_part_sd") := sd(inflations, na.rm = TRUE),
      !!paste0(block_name, "_adj") := mean(inflations[popped == 0], na.rm = TRUE),
      !!paste0(block_name, "_earnings") := sum(inflations[popped == 0], na.rm = TRUE) * 0.003,
      .groups = "drop"
    )
}

# Calculate blocks for each color and bind them
blocks_df <- NULL

for(color_label in colors) {
  for(block_num in 1:6) { # assuming you have up to 6 blocks per color (1-60 trials per color)
    block_stats <- calculate_block_stats(df, color_label, block_num)
    blocks_df <- if(is.null(blocks_df)) block_stats else full_join(blocks_df, block_stats, by = "participant_id")
  }
}

# Join the new data to the original dataframe (repeated values per participant)
df_extended <- left_join(df, blocks_df, by = "participant_id")

# Calculate averages and standard deviations for pre, post, and total blocks per color
blocks_df <- blocks_df %>%
  mutate(
    yellow_pre = rowMeans(select(., starts_with("y") & ends_with("_avg") & matches("y[1-3]_avg")), na.rm = TRUE),
    yellow_post = rowMeans(select(., starts_with("y") & ends_with("_avg") & matches("y[4-6]_avg")), na.rm = TRUE),
    yellow_total = rowMeans(select(., starts_with("y") & ends_with("_avg")), na.rm = TRUE),
    blue_pre = rowMeans(select(., starts_with("b") & ends_with("_avg") & matches("b[1-3]_avg")), na.rm = TRUE),
    blue_post = rowMeans(select(., starts_with("b") & ends_with("_avg") & matches("b[4-6]_avg")), na.rm = TRUE),
    blue_total = rowMeans(select(., starts_with("b") & ends_with("_avg")), na.rm = TRUE),
    orange_pre = rowMeans(select(., starts_with("o") & ends_with("_avg") & matches("o[1-3]_avg")), na.rm = TRUE),
    orange_post = rowMeans(select(., starts_with("o") & ends_with("_avg") & matches("o[4-6]_avg")), na.rm = TRUE),
    orange_total = rowMeans(select(., starts_with("o") & ends_with("_avg")), na.rm = TRUE)
  )

# Join these new averages to the main dataframe
df_extended <- left_join(df_extended, blocks_df %>% select(participant_id, ends_with("_pre"), ends_with("_post"), ends_with("_total")), by = "participant_id")
```

### Group-Level Summary

Calculate the mean and standard deviation from the distribution of block averages across all participants:

```{r}
group_df <- blocks_df %>%
  summarise(across(-participant_id, 
                   list(avg_all = ~mean(.x, na.rm = TRUE), 
                        sd_all = ~sd(.x, na.rm = TRUE))))

# Append group-level data to the original dataframe (as single row repeated)
df_extended <- bind_cols(df_extended, group_df[rep(1, nrow(df_extended)), ])
```






---
title: "BART Block Analysis"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
```

### Data Preparation

Replace `df` with your actual dataframe name.

```{r}
# Assume df is your dataframe
colors <- c("b", "o", "y")
block_size <- 10

# Function to calculate block statistics per color
calculate_block_stats <- function(df, color_label, block_num) {
  block_name <- paste0(color_label, block_num)
  
  df %>%
    group_by(participant_id) %>%
    filter(balloon_color == color_label) %>%
    mutate(color_trial_order = row_number()) %>%
    filter(color_trial_order > (block_num - 1) * block_size &
           color_trial_order <= block_num * block_size) %>%
    summarise(
      !!paste0(block_name, "_avg") := mean(inflations, na.rm = TRUE),
      !!paste0(block_name, "_part_sd") := sd(inflations, na.rm = TRUE),
      !!paste0(block_name, "_adj") := mean(inflations[popped == 0], na.rm = TRUE),
      !!paste0(block_name, "_earnings") := sum(inflations[popped == 0], na.rm = TRUE) * 0.003,
      .groups = "drop"
    )
}

# Calculate blocks for each color and bind them
blocks_df <- NULL

for(color_label in colors) {
  for(block_num in 1:6) { # assuming you have up to 6 blocks per color (1-60 trials per color)
    block_stats <- calculate_block_stats(df, color_label, block_num)
    blocks_df <- if(is.null(blocks_df)) block_stats else full_join(blocks_df, block_stats, by = "participant_id")
  }
}

# Join the new data to the original dataframe (repeated values per participant)
df_extended <- left_join(df, blocks_df, by = "participant_id")

# Calculate averages, adjusted inflations, and earnings for pre, post, and total blocks per color
blocks_df <- blocks_df %>%
  mutate(
    yellow_pre = rowMeans(select(., starts_with("y") & ends_with("_avg") & matches("y[1-3]_avg")), na.rm = TRUE),
    yellow_post = rowMeans(select(., starts_with("y") & ends_with("_avg") & matches("y[4-6]_avg")), na.rm = TRUE),
    yellow_total = rowMeans(select(., starts_with("y") & ends_with("_avg")), na.rm = TRUE),
    blue_pre = rowMeans(select(., starts_with("b") & ends_with("_avg") & matches("b[1-3]_avg")), na.rm = TRUE),
    blue_post = rowMeans(select(., starts_with("b") & ends_with("_avg") & matches("b[4-6]_avg")), na.rm = TRUE),
    blue_total = rowMeans(select(., starts_with("b") & ends_with("_avg")), na.rm = TRUE),
    orange_pre = rowMeans(select(., starts_with("o") & ends_with("_avg") & matches("o[1-3]_avg")), na.rm = TRUE),
    orange_post = rowMeans(select(., starts_with("o") & ends_with("_avg") & matches("o[4-6]_avg")), na.rm = TRUE),
    orange_total = rowMeans(select(., starts_with("o") & ends_with("_avg")), na.rm = TRUE),
    yellow_pre_adj = rowMeans(select(., matches("y[1-3]_adj")), na.rm = TRUE),
    yellow_post_adj = rowMeans(select(., matches("y[4-6]_adj")), na.rm = TRUE),
    yellow_total_adj = rowMeans(select(., matches("y.*_adj")), na.rm = TRUE),
    blue_pre_adj = rowMeans(select(., matches("b[1-3]_adj")), na.rm = TRUE),
    blue_post_adj = rowMeans(select(., matches("b[4-6]_adj")), na.rm = TRUE),
    blue_total_adj = rowMeans(select(., matches("b.*_adj")), na.rm = TRUE),
    orange_pre_adj = rowMeans(select(., matches("o[1-3]_adj")), na.rm = TRUE),
    orange_post_adj = rowMeans(select(., matches("o[4-6]_adj")), na.rm = TRUE),
    orange_total_adj = rowMeans(select(., matches("o.*_adj")), na.rm = TRUE),
    yellow_pre_earnings = rowSums(select(., matches("y[1-3]_earnings")), na.rm = TRUE),
    yellow_post_earnings = rowSums(select(., matches("y[4-6]_earnings")), na.rm = TRUE),
    yellow_total_earnings = rowSums(select(., matches("y.*_earnings")), na.rm = TRUE),
    blue_pre_earnings = rowSums(select(., matches("b[1-3]_earnings")), na.rm = TRUE),
    blue_post_earnings = rowSums(select(., matches("b[4-6]_earnings")), na.rm = TRUE),
    blue_total_earnings = rowSums(select(., matches("b.*_earnings")), na.rm = TRUE),
    orange_pre_earnings = rowSums(select(., matches("o[1-3]_earnings")), na.rm = TRUE),
    orange_post_earnings = rowSums(select(., matches("o[4-6]_earnings")), na.rm = TRUE),
    orange_total_earnings = rowSums(select(., matches("o.*_earnings")), na.rm = TRUE)
  )

# Join these new averages, adjusted inflations, and earnings to the main dataframe
df_extended <- left_join(df_extended, blocks_df %>% select(participant_id, ends_with("_pre"), ends_with("_post"), ends_with("_total")), by = "participant_id")
```

### Group-Level Summary

Calculate the mean and standard deviation from the distribution of block averages and earnings across all participants:

```{r}
group_df <- blocks_df %>%
  summarise(across(-participant_id, 
                   list(avg_all = ~mean(.x, na.rm = TRUE), 
                        sd_all = ~sd(.x, na.rm = TRUE))))

# Append group-level data to the original dataframe (as single row repeated)
df_extended <- bind_cols(df_extended, group_df[rep(1, nrow(df_extended)), ])
```

Your dataframe `df_extended` now contains all required individual and group-level data, and `group_df` contains the summarized group-level data explicitly.






---
title: "BART Block Analysis"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
```

```{r}
# Assume df is your dataframe
colors <- c("b", "o", "y")
block_size <- 10

# Function to calculate block statistics per color
calculate_block_stats <- function(df, color_label, block_num) {
  block_name <- paste0(color_label, block_num)
  
  df %>%
    group_by(participant_id) %>%
    filter(balloon_color == color_label) %>%
    mutate(color_trial_order = row_number()) %>%
    filter(color_trial_order > (block_num - 1) * block_size &
           color_trial_order <= block_num * block_size) %>%
    summarise(
      !!paste0(block_name, "_avg") := mean(inflations, na.rm = TRUE),
      !!paste0(block_name, "_part_sd") := sd(inflations, na.rm = TRUE),
      !!paste0(block_name, "_adj") := mean(inflations[popped == 0], na.rm = TRUE),
      !!paste0(block_name, "_earnings") := sum(inflations[popped == 0], na.rm = TRUE) * 0.003,
      .groups = "drop"
    )
}

# Calculate blocks for each color and bind them
blocks_df <- NULL

for(color_label in colors) {
  for(block_num in 1:6) { # assuming you have up to 6 blocks per color (1-60 trials per color)
    block_stats <- calculate_block_stats(df, color_label, block_num)
    blocks_df <- if(is.null(blocks_df)) block_stats else full_join(blocks_df, block_stats, by = "participant_id")
  }
}

# Join the new data to the original dataframe (repeated values per participant)
df_extended <- left_join(df, blocks_df, by = "participant_id")

# Calculate averages, adjusted inflations, and earnings for pre, post, and total blocks per color
blocks_df <- blocks_df %>%
  mutate(
    yellow_pre = rowMeans(select(., starts_with("y") & ends_with("_avg") & matches("y[1-3]_avg")), na.rm = TRUE),
    yellow_post = rowMeans(select(., starts_with("y") & ends_with("_avg") & matches("y[4-6]_avg")), na.rm = TRUE),
    yellow_total = rowMeans(select(., starts_with("y") & ends_with("_avg")), na.rm = TRUE),
    blue_pre = rowMeans(select(., starts_with("b") & ends_with("_avg") & matches("b[1-3]_avg")), na.rm = TRUE),
    blue_post = rowMeans(select(., starts_with("b") & ends_with("_avg") & matches("b[4-6]_avg")), na.rm = TRUE),
    blue_total = rowMeans(select(., starts_with("b") & ends_with("_avg")), na.rm = TRUE),
    orange_pre = rowMeans(select(., starts_with("o") & ends_with("_avg") & matches("o[1-3]_avg")), na.rm = TRUE),
    orange_post = rowMeans(select(., starts_with("o") & ends_with("_avg") & matches("o[4-6]_avg")), na.rm = TRUE),
    orange_total = rowMeans(select(., starts_with("o") & ends_with("_avg")), na.rm = TRUE),
    yellow_pre_adj = rowMeans(select(., matches("y[1-3]_adj")), na.rm = TRUE),
    yellow_post_adj = rowMeans(select(., matches("y[4-6]_adj")), na.rm = TRUE),
    yellow_total_adj = rowMeans(select(., matches("y.*_adj")), na.rm = TRUE),
    blue_pre_adj = rowMeans(select(., matches("b[1-3]_adj")), na.rm = TRUE),
    blue_post_adj = rowMeans(select(., matches("b[4-6]_adj")), na.rm = TRUE),
    blue_total_adj = rowMeans(select(., matches("b.*_adj")), na.rm = TRUE),
    orange_pre_adj = rowMeans(select(., matches("o[1-3]_adj")), na.rm = TRUE),
    orange_post_adj = rowMeans(select(., matches("o[4-6]_adj")), na.rm = TRUE),
    orange_total_adj = rowMeans(select(., matches("o.*_adj")), na.rm = TRUE),
    yellow_pre_earnings = rowSums(select(., matches("y[1-3]_earnings")), na.rm = TRUE),
    yellow_post_earnings = rowSums(select(., matches("y[4-6]_earnings")), na.rm = TRUE),
    yellow_total_earnings = rowSums(select(., matches("y.*_earnings")), na.rm = TRUE),
    blue_pre_earnings = rowSums(select(., matches("b[1-3]_earnings")), na.rm = TRUE),
    blue_post_earnings = rowSums(select(., matches("b[4-6]_earnings")), na.rm = TRUE),
    blue_total_earnings = rowSums(select(., matches("b.*_earnings")), na.rm = TRUE),
    orange_pre_earnings = rowSums(select(., matches("o[1-3]_earnings")), na.rm = TRUE),
    orange_post_earnings = rowSums(select(., matches("o[4-6]_earnings")), na.rm = TRUE),
    orange_total_earnings = rowSums(select(., matches("o.*_earnings")), na.rm = TRUE)
  )

# Join these new averages, adjusted inflations, and earnings to the main dataframe
df_extended <- left_join(df_extended, blocks_df %>% select(participant_id, ends_with("_pre"), ends_with("_post"), ends_with("_total")), by = "participant_id")
```

### Group-Level Summary

Calculate the mean and standard deviation from the distribution of block averages and earnings across all participants:

```{r}
group_df <- blocks_df %>%
  summarise(across(-participant_id, 
                   list(avg_all = ~mean(.x, na.rm = TRUE), 
                        sd_all = ~sd(.x, na.rm = TRUE))))

# Append group-level data to the original dataframe (as single row repeated)
df_extended <- bind_cols(df_extended, group_df[rep(1, nrow(df_extended)), ])
```
